---
title: "Hands-on Exercise 4"
author: "WAN HONGLU"
date: "Januray 30, 2024"
date-modified: "last-modified"

execute: 
  eval: true
  echo: true
  warning: false
editor: visual
---

# Visualising Distribution
## Loading R packages
```{r}
pacman::p_load(ggdist, ggridges, ggthemes,
               colorspace, tidyverse)
```

## Importing the data
```{r}
exam <- read_csv("data/Exam_data.csv")
```
## Plotting Practice
### Visualising Distribution with Ridgeline Plot

Plotting ridgeline graph: ggridges() method.
```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(exam, 
       aes(x = ENGLISH, 
           y = CLASS)) +
  geom_density_ridges(
    scale = 3,
    rel_min_height = 0.01,
    bandwidth = 3.4,
    fill = lighten("#7097BB", .3),
    color = "white"
  ) +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
    ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

Varying fill colors along the x axis.
```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(exam, 
       aes(x = ENGLISH, 
           y = CLASS,
           fill = stat(x))) +
  geom_density_ridges_gradient(
    scale = 3,
    rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Temp. [F]",
                       option = "C") +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
  ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

Mapping the probabilities directly onto colour.
```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = 0.5 - abs(0.5-stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", 
                      calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail probability",
                       direction = -1) +
  theme_ridges()
```

Ridgeline plots with quantile lines.
```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = 4,
    quantile_lines = TRUE) +
  scale_fill_viridis_d(name = "Quartiles") +
  theme_ridges()
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = c(0.025, 0.975)
    ) +
  scale_fill_manual(
    name = "Probability",
    values = c("#FF0000A0", "#A0A0A0A0", "#0000FFA0"),
    labels = c("(0, 0.025]", "(0.025, 0.975]", "(0.975, 1]")
  ) +
  theme_ridges()
```

### Visualising Distribution with Raincloud Plot
Plotting a Half Eye graph.
```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA)
```

Adding the boxplot with geom_boxplot().
```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA)
```

Adding the Dot Plots with stat_dots().
```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 2)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 1.5) +
  coord_flip() +
  theme_economist()
```

# Visual Statistical Analysis
## Loading R packages_1
```{r}
pacman::p_load(ggstatsplot, tidyverse)
```

## Importing the data_1
```{r}
exam <- read_csv("data/Exam_data.csv")
```
## Plotting Practice_1
### Visual Statistical Analysis with ggstatsplot
One-sample test: gghistostats() method.
```{r}
#| code-fold: true
#| code-summary: "Show the code"
set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```
Two-sample mean test: ggbetweenstats().
```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)
```
Oneway ANOVA Test: ggbetweenstats() method.
```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggbetweenstats(
  data = exam,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```
Significant Test of Correlation: ggscatterstats().
```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggscatterstats(
  data = exam,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )
```
Significant Test of Association (Depedence) : ggbarstats() methods.
```{r}
#| code-fold: true
#| code-summary: "Show the code"
exam1 <- exam %>% 
  mutate(MATHS_bins = 
           cut(MATHS, 
               breaks = c(0,60,75,85,100))
)
ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```
## Loading R packages_2
```{r}
pacman::p_load(readxl, performance, parameters, see)
```

## Importing the data_2
```{r}
car_resale <- read_xls("data/ToyotaCorolla.xls", 
                       "data")
car_resale
```
## Plotting Practice_2
### Visualise model diagnostic and model parameters by using parameters package
Multiple Regression Model using lm().
```{r}
model <- lm(Price ~ Age_08_04 + Mfg_Year + KM + 
              Weight + Guarantee_Period, data = car_resale)
model
```

Model Diagnostic: checking for multicolinearity.
```{r}
check_collinearity(model)
```
```{r}
check_c <- check_collinearity(model)
plot(check_c)
```

Model Diagnostic: checking normality assumption.
```{r}
#| code-fold: true
#| code-summary: "Show the code"
model1 <- lm(Price ~ Age_08_04 + KM + 
              Weight + Guarantee_Period, data = car_resale)

check_n <- check_normality(model1)

plot(check_n)
```
Model Diagnostic: Check model for homogeneity of variances.
```{r}
#| code-fold: true
#| code-summary: "Show the code"
check_h <- check_heteroscedasticity(model1)

plot(check_h)
```
Model Diagnostic: Complete check.
```{r}
check_model(model1)
```
Visualising Regression Parameters: see methods.
```{r}
plot(parameters(model1))
```

Visualising Regression Parameters: ggcoefstats() methods.
```{r}
ggcoefstats(model1, 
            output = "plot")
```



