---
title: "Hands-on Exercise 6 :Visualising and Analysing Time-oriented Data"
author: "WAN HONGLU"
date: "Feb 12, 2024"
date-modified: "last-modified"

execute:
  eval: true
  echo: true
  warning: false
editor: visual
---

## Install and Load packages

```{r}
pacman::p_load(tidyverse, readxl, data.table,
               ggthemes, scales, viridis, 
               knitr, lubridate, gridExtra, CGPfunctions)
```
## Ploting Practice - Plotting Calendar Heatmap
### Importing the data

```{r}
attacks <- read_csv("data/eventlog.csv")

kable(head(attacks))
```

### Data Preparation
**Step 1 : Deriving weekday and hour of day fields**

```{r}
#| code-fold: true
#| code-summary: "Show the code"
make_hr_wkday <- function(ts, sc, tz) {
  real_times <- ymd_hms(ts, 
                        tz = tz[1], 
                        quiet = TRUE)
  dt <- data.table(source_country = sc,
                   wkday = weekdays(real_times),
                   hour = hour(real_times))
  return(dt)
}
```

**Step 2: Deriving the attacks tibble data frame**

```{r}
#| code-fold: true
#| code-summary: "Show the code"
wkday_levels <- c('Saturday', 'Friday', 
                  'Thursday', 'Wednesday', 
                  'Tuesday', 'Monday', 
                  'Sunday')

attacks <- attacks %>%
  group_by(tz) %>%
  do(make_hr_wkday(.$timestamp, 
                   .$source_country, 
                   .$tz)) %>% 
  ungroup() %>% 
  mutate(wkday = factor(
    wkday, levels = wkday_levels),
    hour  = factor(
      hour, levels = 0:23))

kable(head(attacks))
```

### Building the Calendar Heatmaps
```{r}
#| code-fold: true
#| code-summary: "Show the code"
grouped <- attacks %>% 
  count(wkday, hour) %>% 
  ungroup() %>%
  na.omit()

ggplot(grouped, 
       aes(hour, 
           wkday, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
theme_tufte(base_family = "Helvetica") + 
coord_equal() +
scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = NULL, 
     y = NULL, 
     title = "Attacks by weekday and time of day") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6) )
```

### Building Multiple Calendar Heatmaps
**Step 1:** Deriving attack by country object.

**Step 2:** Preparing the tidy data frame.

**Step 3:** Plotting the Multiple Calender Heatmap by using ggplot2 package.
```{r}
#| code-fold: true
#| code-summary: "Show the code"
attacks_by_country <- count(
  attacks, source_country) %>%
  mutate(percent = percent(n/sum(n))) %>%
  arrange(desc(n))

top4 <- attacks_by_country$source_country[1:4]
top4_attacks <- attacks %>%
  filter(source_country %in% top4) %>%
  count(source_country, wkday, hour) %>%
  ungroup() %>%
  mutate(source_country = factor(
    source_country, levels = top4)) %>%
  na.omit()

ggplot(top4_attacks, 
       aes(hour, 
           wkday, 
           fill = n)) + 
  geom_tile(color = "white", 
          size = 0.1) + 
  theme_tufte(base_family = "Helvetica") + 
  coord_equal() +
  scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
  facet_wrap(~source_country, ncol = 2) +
  labs(x = NULL, y = NULL, 
     title = "Attacks on top 4 countries by weekday and time of day") +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6) )
```

## Plotting Practice - Plotting Cycle Plot
### Step 1: Data Import
```{r}
air <- read_excel("data/arrivals_by_air.xlsx")
```

### Step 2: Deriving month and year fields
```{r}
#| code-fold: true
#| code-summary: "Show the code"
air$month <- factor(month(air$`Month-Year`), 
                    levels=1:12, 
                    labels=month.abb, 
                    ordered=TRUE) 
air$year <- year(ymd(air$`Month-Year`))
```

### Step 3: Extracting the target country
```{r}
#| code-fold: true
#| code-summary: "Show the code"
Vietnam <- air %>% 
  select(`Vietnam`, 
         month, 
         year) %>%
  filter(year >= 2010)
```

### Step 4: Computing year average arrivals by month
```{r}
#| code-fold: true
#| code-summary: "Show the code"
hline.data <- Vietnam %>% 
  group_by(month) %>%
  summarise(avgvalue = mean(`Vietnam`))
```

### Step 5: Plotting the cycle plot
```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot() + 
  geom_line(data=Vietnam,
            aes(x=year, 
                y=`Vietnam`, 
                group=month), 
            colour="black") +
  geom_hline(aes(yintercept=avgvalue), 
             data=hline.data, 
             linetype=6, 
             colour="red", 
             size=0.5) + 
  facet_grid(~month) +
  labs(axis.text.x = element_blank(),
       title = "Visitor arrivals from Vietnam by air, Jan 2010-Dec 2019") +
  xlab("") +
  ylab("No. of Visitors") +
  theme_tufte(base_family = "Helvetica")
```

## Plotting Practice - Plotting Slopegraph
### Step 1: Data Import
```{r}
rice <- read_csv("data/rice.csv")
```

### Step 2: Plotting the slopegraph
```{r}
#| code-fold: true
#| code-summary: "Show the code"
rice %>% 
  mutate(Year = factor(Year)) %>%
  filter(Year %in% c(1961, 1980)) %>%
  newggslopegraph(Year, Yield, Country,
                Title = "Rice Yield of Top 11 Asian Counties",
                SubTitle = "1961-1980",
                Caption = "Prepared by: Dr. Kam Tin Seong")
```

