---
title: "Hands-on Exercise 8: Visualising and Analysing Geographic Data"
author: "WAN HONGLU"
date: "Feb 24, 2024"
date-modified: "last-modified"

execute:
  eval: true
  echo: true
  warning: false
editor: visual
---

# Choropleth Mapping with R

## Install and Load packages

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

## Importing Data

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
mpsz
```

## Importing Attribute Data

```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

## Data Preparation

### Data wrangling

```{r}
#| code-fold: true
#| code-summary: "Show the code"
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```

### Joining the attribute data and geospatial data

```{r}
#| code-fold: true
#| code-summary: "Show the code"
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)

mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))

write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

## Plotting Practice - Choropleth Mapping Geospatial Data Using tmap

### Plotting a choropleth map quickly by using qtm()

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

### Creating a choropleth map by using tmap’s elements

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

In the following sub-section to share tmap functions that used to plot these elements. Step 1: Drawing a base map.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

Step 2: Drawing a choropleth map using tm_polygons().

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

Step 3: Drawing a choropleth map using tm_fill() and \*tm_border().

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,  alpha = 1)
```

### Data classification methods of tmap

Plotting choropleth maps with built-in classification methods.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```

### Plotting choropleth map with custome break

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```

### Colour Scheme

Using ColourBrewer palette.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```

### Map Layouts

Map Legend.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

Map style.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

Cartographic Furniture.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))

```

```{r}
tmap_style("white")
```

### Drawing Small Multiple Choropleth Maps

By assigning multiple values to at least one of the aesthetic arguments.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

By defining a group-by variable in tm_facets().

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE, 
            drop.shapes=FALSE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

By creating multiple stand-alone maps with tmap_arrange().

```{r}
#| code-fold: true
#| code-summary: "Show the code"
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

### Mappping Spatial Object Meeting a Selection Criterion

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

# Visualising Geospatial Point Data

## Install and Load packages

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

## Data Import and Preparation

```{r}
sgpools <- read_csv("data/aspatial/SGPools_svy21.csv")
```

```{r}
list(sgpools) 
```

## Plotting Practice - Creating a sf data frame from an aspatial data frame

```{r}
#| code-fold: true
#| code-summary: "Show the code"
sgpools_sf <- st_as_sf(sgpools, 
                       coords = c("XCOORD", "YCOORD"),
                       crs= 3414)
list(sgpools_sf)
```

## Plotting Practice - Drawing Proportional Symbol Map

```{r}
tmap_mode("view")
```

### It all started with an interactive point symbol map

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(sgpools_sf)+
tm_bubbles(col = "red",
           size = 1,
           border.col = "black",
           border.lwd = 1)
```

### Making it proportional

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(sgpools_sf)+
tm_bubbles(col = "red",
           size = "Gp1Gp2 Winnings",
           border.col = "black",
           border.lwd = 1)
```

### Giving it a different colour

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(sgpools_sf)+
tm_bubbles(col = "OUTLET TYPE", 
          size = "Gp1Gp2 Winnings",
          border.col = "black",
          border.lwd = 1)
```

### Having a twin brothers

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(sgpools_sf) +
  tm_bubbles(col = "OUTLET TYPE", 
          size = "Gp1Gp2 Winnings",
          border.col = "black",
          border.lwd = 1) +
  tm_facets(by= "OUTLET TYPE",
            nrow = 1,
            sync = TRUE)
```

```{r}
tmap_mode("plot")
```

# Analytical Mapping

## Installing and loading packages

```{r}
pacman::p_load(tmap, tidyverse, sf)
```

## Importing data

```{r}
NGA_wp <- read_rds("data/rds/NGA_wp.rds")
```

## Plotting Practice - Basic Choropleth Mapping

### Visualising distribution of non-functional water point

```{r}
#| code-fold: true
#| code-summary: "Show the code"
p1 <- tm_shape(NGA_wp) +
  tm_fill("wp_functional",
          n = 10,
          style = "equal",
          palette = "Blues") +
  tm_borders(lwd = 0.1,
             alpha = 1) +
  tm_layout(main.title = "Distribution of functional water point by LGAs",
            legend.outside = FALSE)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
p2 <- tm_shape(NGA_wp) +
  tm_fill("total_wp",
          n = 10,
          style = "equal",
          palette = "Blues") +
  tm_borders(lwd = 0.1,
             alpha = 1) +
  tm_layout(main.title = "Distribution of total  water point by LGAs",
            legend.outside = FALSE)
```

```{r}
tmap_arrange(p2, p1, nrow = 1)
```

## Plotting Practice - Choropleth Map for Rates

### Deriving Proportion of Functional Water Points and Non-Functional Water Points

```{r}
#| code-fold: true
#| code-summary: "Show the code"
NGA_wp <- NGA_wp %>%
  mutate(pct_functional = wp_functional/total_wp) %>%
  mutate(pct_nonfunctional = wp_nonfunctional/total_wp)
```

### Plotting map of rate

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(NGA_wp) +
  tm_fill("pct_functional",
          n = 10,
          style = "equal",
          palette = "Blues",
          legend.hist = TRUE) +
  tm_borders(lwd = 0.1,
             alpha = 1) +
  tm_layout(main.title = "Rate map of functional water point by LGAs",
            legend.outside = TRUE)
```

## Plotting Practice - Extreme Value Maps

### Percentile Map

**Data Preparation.**

Step 1: Exclude records with NA by using the code chunk below.

```{r}
NGA_wp <- NGA_wp %>%
  drop_na()
```

Step 2: Creating customised classification and extracting values.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
percent <- c(0,.01,.1,.5,.9,.99,1)
var <- NGA_wp["pct_functional"] %>%
  st_set_geometry(NULL)
quantile(var[,1], percent)
```

Creating the get.var function.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
get.var <- function(vname,df) {
  v <- df[vname] %>% 
    st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

A percentile mapping function.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var <- get.var(vnam, df)
  bperc <- quantile(var, percent)
  tm_shape(df) +
  tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,
             title=legtitle,
             breaks=bperc,
             palette="Blues",
          labels=c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("right","bottom"))
}
```

Test drive the percentile mapping function.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
percentmap("total_wp", NGA_wp)
```

### Box map

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggplot(data = NGA_wp,
       aes(x = "",
           y = wp_nonfunctional)) +
  geom_boxplot()
```

Creating the boxbreaks function.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```

Creating the get.var function.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

Test drive the newly created function.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
var <- get.var("wp_nonfunctional", NGA_wp) 
boxbreaks(var)
```

Boxmap function.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
    tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("left",
                               "top"))
}
tmap_mode("plot")
boxmap("wp_nonfunctional", NGA_wp)
```
